The Galaxy Clustering Crisis in Abundance Matching

### 2017-05-18 paper/paper.tex
lots of updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
%\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{epsfig}
\usepackage{hyperref}
\hypersetup{draft}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

\usepackage{lineno}
\linenumbers

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\usepackage{astrobib_mnras2e}
\input{macros_dc.sty}
\providecommand{\eprint}[1]{\href{http://arxiv.org/abs/#1}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title[The Galaxy Clustering Crisis]{The Galaxy Clustering Crisis in Abundance Matching}
\author[Campbell et al.]
{\parbox[t]{\textwidth}{
Duncan Campbell$^{1}$\thanks{E-mail: duncan.campbell@yale.edu},
Frank C. van den Bosch$^{1}$,
Nikhil Padmanabhan$^{2, 3}$,\\ 
Johannes Lange$^{1}$,
Yao-Yuan Mao$^{4}$, 
Fangzhou Jiang$^{5}$,
Antonio Villarreal$^{4}$,\\
and Andrew Zentner$^{4}$
} \\
\vspace*{3pt} \\
$^{1}$Department of Astronomy, Yale University, New Haven, CT 06511, USA\\
$^{2}$Department of Physics, Yale University, New Haven, CT 06520, USA \\
$^{3}$Yale Center for Astronomy and Astrophysics, Yale University, New Haven, CT 06520, USA\\
$^{4}$Department of Physics and Astronomy \& Pittsburgh Particle Physics, Astrophysics, and Cosmology Center (PITT PACC),\\ University of Pittsburgh, Pittsburgh, PA 15260\\
$^{5}$Racah Institute of Physics, The Hebrew University, Jerusalem 91904, Israel
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2016}

@@

\begin{abstract}
Galaxy clustering on small scales is significantly under-predicted by sub-halo abundance matching (SHAM) models that populate (sub-)haloes with galaxies based on peak halo mass, $M_{\rm peak}$.  SHAM methods based on the peak maximum circular velocity, $V_{\rm peak}$, have had much better success.  The primary reason $M_{\rm peak}$ based models fail is the relatively low abundance of satellite galaxies produced in these models compared to those based on $V_{\rm peak}$.  Despite success in predicting clustering, a simple $V_{\rm peak}$ based SHAM model results in predictions for galaxy growth that are at odds with observations.  We evaluate three possible remedies that could ``save'' mass-based SHAM.  1.) SHAM models require a significant population of ``orphan'' galaxies as a result of artificial disruption/merging of sub-haloes in modern high resolution dark matter simulations, 2.) satellites must grow significantly after being accretion, and/or 3.) stellar mass is significantly affected by halo assembly history.  Regardless of the particulars, we show that popular SHAM models based on $M_{\rm peak}$ cannot be correct as presented.  Either $V_{\rm peak}$ truly is a better predictor of stellar mass, or SHAM models are missing vital component(s) that significantly affect galaxy clustering.
\end{abstract}

@@

The prevailing picture of galaxy formation is intricately tied to that of dark matter structure formation.  The matter distribution of the Universe evolved from a relatively smooth state into a complex web of structure over $\sim 14$ billion years.  Within this web, small inhomogeneities evolved into extended, quasi-spherical, gravitationally bound objects called haloes.  The build-up of structure proceeds hierarchically as smaller haloes are accreted into larger haloes, becoming substructures called sub-haloes.  The potential wells of dark matter (sub-)haloes are the natural site for galaxies to form as baryons cool and condense into stars \citep{Rees:1977gr, White:1978uk, Fall:1980br}.

This general, albeit coarse, view of galaxy formation fits well with the premise of sub-halo abundance matching (SHAM).  SHAM in its most simple form rests on the hypothesis that all massive\footnote{above a limiting lower mass scale below which galaxy formation becomes inefficient} (sub-)haloes host galaxies, all galaxies occupy (sub-)haloes, and there is a simple monotonic relation between galaxy mass and the mass of the (sub-)halo each galaxy occupies \citep{Kravtsov:2004fi, Conroy:2006iz, Vale:2004bb}.  This approach relies on simulations of dark matter structure formation to provide robust statistical predictions for many properties of dark matter (sub-)haloes, e.g. halo mass functions, mass profiles, and the spatial clustering of haloes.  By ``populating'' dark matter simulations with galaxies using the SHAM technique, the statistical predictions from simulations can be leveraged to constrain the galaxy-halo connection and make predictions for how galaxies form and evolve along with (sub-)haloes.

The stellar mass-halo mass (SMHM) relation is one of the most fundamental implications of SHAM.  The SMHM relation at redshift $\sim 0$ inferred using SHAM is consistent with HOD/CLF analysis \citep{Yang:2003du, vandenBosch:2003fd, vandenBosch:2007ea, Yang:2013hw, Zu:2015tt} and independent, more direct, measurements,  e.g. galaxy-galaxy lensing, satellite kinematics, and the Tully-Fisher relation \citep{Wang:2006fe, More:2009ws, Guo:2010do, Wang:2010ho, Moster:2010ep, Behroozi:2010ja, Mandelbaum:2015wt, Desmond:2015gr, vanUitert:2016fy}.  By extending this analysis to higher redshifts, the inferred evolution of the SMHM relation and the mass growth histories of haloes predicted by simulations provide constraints on the average stellar mass growth histories of galaxies that is consistent with the cosmic star-formation history of the Universe \citep{Conroy:2009ba, Wang:2010ho, Yang:2012ew, Moster:2013ab, Behroozi:2013fg}.  An important finding of these studies is that the peak in star-formation efficiency shifts towards more massive haloes at earlier times.

The ability of SHAM to accurately predict the clustering of galaxies is more limited.  In particular, the small scale clustering of galaxies remains difficult to fit.  SHAM has only been shown to be consistent with galaxy clustering under one of two conditions: 1.) stellar mass is tied to the peak circular velocity of (sub-)haloes (or closely related quantities) \citep{Reddick:2013gi, Hearin:2014hh, Lehmann:2015vi}, or 2.) the abundance of sub-haloes is treated as a free parameter \citep{Wang:2006fe, Wang:2010ho, Guo:2010do, Moster:2010ep, Yang:2012ew}.  This second approach relies on a population of ``orphan'' galaxies which have no identifiable sub-halo in a simulation.  As we will show in this work, both of these assumptions are problematic.  The assumption that stellar mass should be better correlated with maximum circular velocity over halo mass has not been robustly motivated.  Furthermore, we show that models that make this assumption are at odds with previous work on the evolution of the SMHM relation.  Orphan galaxies are difficult to reconcile with modern high resolution dark matter simulations which aim to resolve substructure.  We summarize this state of affairs in Fig. \ref{fig:venn_diagram}.  There is no published SHAM model which fits both galaxy clustering and the evolution of the stellar mass function using resolved substructure in a dark matter simulation without the addition of ``orphan'' galaxies.

\begin{figure}
    \includegraphics{figures/venn_diagram.pdf}
    \caption{Venn diagram of selected SHAM models within three sets; those that fit the galaxy clustering, $w_p(r_p)$ (upper left red region), those that fit the evolution of the galaxy stellar mass function, $\phi(M_*,z)$ (upper right blue region), and those that utilize orphan galaxies (lower green region).  Models marked with a circle are based on measures of halo mass, $M_{\rm vir}$, including peak quantities.  Models marked with a star are based on measures of halo maximum circular velocity, $V_{\rm max}$, including peak quantities, with the \citet{Lehmann:2015vi} model falling somewhere in between.   This paper focuses on the lack of models in the region marked with a ``?''. }
    \label{fig:venn_diagram}
\end{figure}

The goal of this paper is threefold: to make explicit the tension between fitting galaxy clustering measurements and the evolution of the stellar mass function within the SHAM framework, to examine the successes and failures of SHAM models based on maximum circular velocity, and to evaluate mechanisms to alleviate the clustering crisis in models based on halo mass.  The models and analysis in this paper utilize the framework and and code base in {\tt Halotools} \citep{Hearin:2016tc}, an {\tt Astropy} \citep{TheAstropyCollaboration:2013cd} affiliated Python\footnote{http://www.python.org} package.  We also make available all of the code and data products necessary to reproduce the figures and analysis in this paper\footnote{https://github.com/duncandc/Clustering\_Crisis}.  Throughout, we scale all units by $h = H_0/[100~{\rm km/s}~{\rm Mpc}^{-1}]$ where appropriate, and we use $\log(x)$ to indicate the base 10 logarithm of $x$.

\section{Methods}

@@

SHAM, in its most simple form, assumes that the cumulative abundance of galaxies and haloes can be used to map galaxy properties uniquely onto (sub-)haloes.  Using stellar mass, $M_{*}$, and a halo mass proxy $\mathcal{M}$, SHAM assumes:
%
\begin{linenomath}
\begin{equation}
\label{eq:sham}
N(>M_*) = N(>\mathcal{M}) \implies M_{*} = f(\mathcal{M})
\end{equation} 
\end{linenomath}
%
where $f(\mathcal{M})$ is some monotonically increasing function.  This maps the most massive galaxies to the most massive haloes.  The function $f(\mathcal{M})$ may be determined non-parametrically by matching the rank orders of galaxies and haloes.  We refer to this method of SHAM as ``rank order'' abundance matching.  However, we also consider another class of SHAM models where $f(\mathcal{M})$ is parametrized.  This introduces the possibility that the observed abundance of galaxies is not strictly matched; however, it is generally the goal of such models to preserve this characteristic.  We consider these two types of models similar enough to be contained within the class of models we refer to as SHAM.

@@

The simplest implementation of SHAM maps galaxies into (sub-)haloes by matching rank orders.  Given a set of galaxies, $M_*=\{M_{*i}\}$, and (sub-)haloes, $\mathcal{M}=\{\mathcal{M}_j\}$, of the same size, this method proceeds by calculating the ranks of each:
%
\begin{linenomath}
\begin{align}
n^{\rm gal} &= \mathcal{R}(M_*) \nonumber \\
n^{\rm halo} &= \mathcal{R}(\mathcal{M})
\end{align}
\end{linenomath}
%
where the $\mathcal{R}()$ function returns the ordinal ranks.  Each galaxy is then assigned to a (sub-)halo with the equivalent rank, i.e. $n_i^{\rm gal}=n_j^{\rm halo}$.  We consider two rank order SHAM models in this paper, one based on $M_{\rm peak}$ and one on $V_{\rm peak}$.  As shorthand, we refer to these models as ``RM'' and ``RV'', respectively.

Given a set of galaxies and haloes, this mapping is deterministic.  However, the assumption of perfect rank ordering with respect to the cumulative distribution can be relaxed, allowing for stochasticity in the mapping and resulting in a probabilistic relation between stellar mass and halo mass, $P(M_*|\mathcal{M})$.  There are various methods used in the literature to add scatter to this relation while maintaining agreement with an observed stellar mass function.  One may attempt to deconvolve the stellar mass function from the scatter model such that, after solving for eq. (\ref{eq:sham}) with this deconvolved function and applying the scatter model, the new stellar mass function is consistent with the observed function \citep[e.g.][]{Behroozi:2010ja}.  Another option is to manually add scatter to the stellar masses used in abundance matching, re-ranking on the new values to perform the matching, and iteratively solving for a solution that results in the desired amount of scatter in the SMHM relation \citep[e.g.][]{Hearin:2013ok}.  A third option is to parametrize the SMHM relation and fit for the parameters which result in a stellar mass function that is consistent with an observed function after populating a dark matter simulation.  We will examine this type of model in the following sections.

We do not go through the additional step of adding scatter to our RM and RV models, as scatter in the SMHM relation generally decreases the strength of the clustering signal for massive galaxies ($M* > 10^{11} ~h^{-2} M_{\odot}$) with little effect at lower masses \citep{Tinker:2016vu}, and this work focuses on the decrement in galaxy clustering signal for samples of less massive galaxies in SHAM models.

%Many halo properties have been explored in a search for the optimum $\mathcal{M}$ which most faithfully reproduces observed galaxy clustering statistics.  Instantaneous (sub-)halo mass drastically under-predicts clustering \citep{Conroy:2006iz, Reddick:2013gi}, because sub-haloes lose mass via stripping, while the galaxies at their cores are more resilient, resulting in too little power on small scales ($< 1$ Mpc).  Peak values of $\mathcal{M}$, the maximum value reached as a (sub-)halo evolves, have been found to be more successful at reproducing the projected correlation function, $w_p$, in SDSS.  Specifically, \cite{Reddick:2013gi} find that $V_{\rm peak}$ is the most successful in this respect.

@@

\cite{Moster:2013ab} parametrize the SMHM relation as a function of halo mass and redshift to account for evolution.  Furthermore, they assume the stellar mass of satellite galaxies is determined by the mass of the sub-halo at the time of its accretion into a more massive host-halo for the first time.  The functional form is given by:
%
\begin{linenomath}
\begin{equation}
\label{eq:moster_smhm}
\frac{\langle M_{*} | \mathcal{M}\rangle_{\rm med}(a)}{\mathcal{M}} = 2N\left[ \left(\frac{\mathcal{M}}{{M}_1} \right)^{-\beta} + \left( \frac{\mathcal{M}}{{M}_1} \right)^{\gamma} \right]^{-1}
\end{equation}
\end{linenomath}
%
where $\langle\rangle_{\rm med}$ indicates the median.  The evolution of the parameters is given by:
%
\begin{linenomath}
\begin{align}
\label{eq:smhm_params}
\log[{M}_1(a)]& = {M}_{10} + {M}_{11}(1-a) \\
N(a) &= N_{10} + N_{11}(1-a) \\
\beta(a) &= \beta_{10} + \beta_{11}(1-a) \\ 
\gamma(a) &=  \gamma_{10} + \gamma_{11}(1-a)
\end{align}
\end{linenomath}
%
where the scale factor is either the instantaneous scale factor, or the one at the time of accretion:
%
\begin{linenomath}
\begin{equation}
a = 
\begin{cases}
    a      & \quad \text{if host-halo}\\
    a_{\rm acc}  & \quad \text{if sub-halo} \\
  \end{cases}
\end{equation}
\end{linenomath}
%
and $\mathcal{M}$ is the instantaneous mass or the mass at accretion, $M_{\rm acc}$, for sub-haloes.  Scatter in the SMHM relation is modelled as a fixed log-normal with $\sigma_{\log(M*)} \approx 0.18$.

\citet{Moster:2013ab} find the best fit parameters for equation (\ref{eq:moster_smhm}) that reproduce the stellar mass function at various redshifts between 0 and $\sim 4$.  They also show that the implied star-formation rates of galaxies given the growth history of the (sub-)haloes in their simulation is consistent with the cosmic star-formation rate density.  The parameters were constrained using the Millennium simulation with (sub-)haloes defined as spherical over-densities (SO) with mean internal density 200 times the critical density of the universe, $M_{200c}$.   As shorthand, we will refer to this model as ``M13''.  The parameter values used in this paper are taken directly from \citet{Moster:2013ab} and listed in table 2.

@@

\begin{figure}
    \includegraphics{figures/stellar_mass_function.pdf}
    \caption{The stellar mass function of a mock realization of each model is plotted (lines).  For comparison we show the stellar mass function in SDSS as measured by \citet{Li:2009kh} as points with error bars.  The lower panel shows the residuals with respect to the triple Schecter fit from \citet{Li:2009kh}.  We show the $\pm 1 \sigma$ error on the model prediction for the RV as the shaded magenta region model as an example {\bf [DC: still adjusting this to bring each model into better agreement]}}
    \label{fig:phi}
\end{figure}

>>>>

\documentclass[useAMS,fleqn,usenatbib]{mnras}
\pdfoutput=1 %for arxiv
\setlength{\topmargin}{-0.3in}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath,amssymb,amstext}
\usepackage[T1]{fontenc}
\usepackage{newtxtext,newtxmath}
\usepackage[figure,figure*]{hypcap}
\usepackage[most]{tcolorbox}
\input{hyperlink-year-only-natbib-patch}

\tcbset{
    frame code={}
    center title,
    left=0pt,
    right=0pt,
    top=0pt,
    bottom=0pt,
    colback=gray!70,
    colframe=white,
    width=42pt,
    enlarge left by=0mm,
    boxsep=0pt,
    arc=0pt,outer arc=0pt,
    }

%\usepackage{lineno}
%\linenumbers
\usepackage[normalem]{ulem}

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\input{macros_dc.sty}

% for comments:
\newcommand{\dac}[1]{\textcolor{cyan}{\textbf{ #1}}}
\newcommand*{\yao}[1]{\textcolor{magenta}{\textbf{[YYM: #1]}}}
\newcommand{\arz}[1]{\textcolor{red}{\textbf{[ARZ: #1]}}}
\definecolor{BurntOrange}{RGB}{207, 83, 0}
\newcommand{\fvdb}[1]{\textcolor{BurntOrange}{\textbf{[FvdB: #1]}}}
\newcommand{\asv}[1]{\textcolor{purple}{\textbf{[ASV: #1]}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title[The Galaxy Clustering Crisis]{The Galaxy Clustering Crisis in Abundance Matching}
\author[Campbell et al.]{%
Duncan Campbell$^{1}$\thanks{E-mail:  \mailto{duncan.campbell@yale.edu}},
Frank C. van den Bosch$^{1}$,
Nikhil Padmanabhan$^{2, 3}$,\newauthor 
Yao-Yuan Mao$^{4}$, 
Andrew R. Zentner$^{4}$,
Johannes U. Lange$^{1, 5}$,
Fangzhou Jiang$^{6}$,\newauthor 
and Antonio Villarreal$^{4}$\vspace*{0.2em} \\ 
$^{1}$Department of Astronomy, Yale University, New Haven, CT 06511, USA\\
$^{2}$Department of Physics, Yale University, New Haven, CT 06520, USA \\
$^{3}$Yale Center for Astronomy and Astrophysics, Yale University, New Haven, CT 06520, USA\\
$^{4}$Department of Physics and Astronomy \& Pittsburgh Particle Physics, Astrophysics, and Cosmology Center (PITT PACC),\\ University of Pittsburgh, Pittsburgh, PA 15260, USA\\
$^{5}$Kavli Institute for Theoretical Physics, University of California, Santa Barbara, CA 93106, USA\\
$^{6}$Racah Institute of Physics, The Hebrew University, Jerusalem 91904, Israel
}

% These dates will be filled out by the publisher
% \date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2017}

@@

\begin{abstract}
Galaxy clustering on small scales is significantly under-predicted by sub-halo abundance matching (SHAM) models that populate (sub-)haloes with galaxies based on peak halo mass, $M_{\rm peak}$.  SHAM models based on the peak maximum circular velocity, $V_{\rm peak}$, have had much better success.  The primary reason $M_{\rm peak}$ based models fail is the relatively low abundance of satellite galaxies produced in these models compared to those based on $V_{\rm peak}$.  Despite success in predicting clustering, a simple $V_{\rm peak}$ based SHAM model results in predictions for galaxy growth that are at odds with observations.  We evaluate three possible remedies that could ``save'' mass-based SHAM: (1) SHAM models require a significant population of ``orphan'' galaxies as a result of artificial disruption/merging of sub-haloes in modern high resolution dark matter simulations; (2) satellites must grow significantly after their accretion; and (3) stellar mass is significantly affected by halo assembly history. No solution is entirely satisfactory. However, regardless of the particulars, we show that popular SHAM models based on $M_{\rm peak}$ cannot be complete physical models as presented.  Either $V_{\rm peak}$ truly is a better predictor of stellar mass at $z\sim 0$ and it remains to be seen how the correlation between stellar mass and $V_{\rm peak}$ comes about, or SHAM models are missing vital component(s) that significantly affect galaxy clustering.
\end{abstract}

@@

The prevailing picture of galaxy formation is intricately tied to that of dark matter structure formation.  The matter distribution of the Universe evolved from a relatively smooth state into a complex web of structure over $\sim 14$ billion years. Within this web, small inhomogeneities evolved into extended, quasi-spherical, gravitationally bound objects called haloes. The build-up of structure proceeds hierarchically as smaller haloes are accreted into larger haloes, becoming substructures called sub-haloes. The potential wells of dark matter (sub-)haloes are the natural sites for galaxies to form as baryons cool and condense into stars \citep{Rees:1977gr, White:1978uk, Fall:1980br}.

This general, albeit coarse, view of galaxy formation fits well with the premise of sub-halo abundance matching (SHAM).  SHAM in its most simple form rests on the hypothesis that all massive\footnote{above a limiting lower mass scale below which galaxy formation becomes inefficient} (sub-)haloes host galaxies, all galaxies occupy (sub-)haloes, and there is a simple monotonic relation between galaxy mass and the mass of the (sub-)halo each galaxy occupies \citep{Kravtsov:2004fi, Vale:2004bb, Conroy:2006iz}.  This approach relies on simulations of dark matter structure formation to provide robust statistical predictions for many properties of dark matter (sub-)haloes, e.g. halo mass functions, mass profiles, and the spatial clustering of haloes.  By ``populating'' dark matter simulations with galaxies using the SHAM technique, the statistical predictions from simulations can be leveraged to constrain the galaxy-halo connection and make predictions for how galaxies form and evolve along with (sub-)haloes \citep{Springel:2006fq}.

The stellar mass-halo mass (SMHM) relation is one of the most fundamental implications of SHAM.  The SMHM relation at redshift $\sim 0$ inferred using SHAM is consistent with HOD/CLF analysis \citep{Yang:2003du, vandenBosch:2003fd, vandenBosch:2007ea, Yang:2013hw, Zu:2015tt} and with independent, more direct, measurements from galaxy--galaxy lensing, satellite kinematics, and the Tully--Fisher relation \citep{vandenBosch:2004mn, Wang:2006fe, More:2009ws, Guo:2010do, Wang:2010ho, Moster:2010ep, Behroozi:2010ja, Mandelbaum:2015wt, Desmond:2015gr, vanUitert:2016fy}.  By extending this analysis to higher redshifts, the inferred evolution of the SMHM relation and the mass growth histories of haloes predicted by simulations provide constraints on the average stellar mass growth histories of galaxies that are consistent with the cosmic star-formation history of the Universe \citep{Conroy:2009ba, Wang:2010ho, Yang:2012ew, Moster:2013ab, Behroozi:2013fg, RodriguezPuebla:2017uo}.  An important finding of these studies is that the peak in star-formation efficiency shifts towards more massive haloes at earlier times.

The ability of SHAM to accurately predict the clustering of galaxies is more limited.  In particular, the small scale clustering of galaxies remains difficult to fit.  SHAM has only been shown to be consistent with galaxy clustering under one of two conditions: either stellar mass is tied to the peak circular velocity of (sub-)haloes (or closely related quantities) \citep{Reddick:2013gi, Hearin:2014hh, Lehmann:2017fy}, or the abundance of sub-haloes is treated as a free parameter \citep{Wang:2006fe, Wang:2010ho, Guo:2010do, Moster:2010ep, Yang:2012ew}.  This second approach relies on a population of ``orphan'' galaxies which have no identifiable sub-halo in a simulation.  As we will show in this work, both of these assumptions are problematic.  The assumption that stellar mass should be better correlated with maximum circular velocity over halo mass has not been robustly motivated.  Furthermore, we show that models that make this assumption are at odds with previous work on the evolution of the SMHM relation.  Orphan galaxies are difficult to reconcile with modern high resolution dark matter simulations which aim to resolve substructure.  We summarize this state of affairs in Fig. \ref{fig:venn_diagram}.  There is no published SHAM model which fits both galaxy clustering and the evolution of the stellar mass function using resolved substructure in a dark matter simulation without the addition of ``orphan'' galaxies.

\begin{figure}
    \includegraphics[width=\columnwidth]{figures/SHAM_models_venn_diagram.pdf}
    \caption{Venn diagram of selected SHAM models within three sets; those that fit the galaxy clustering, $w_p(r_p)$ (upper left red region), those that fit the evolution of the galaxy stellar mass function, $\phi(M_*,z)$ (upper right blue region), and those that utilize orphan galaxies (lower green region).  Models marked with a circle are based on measures of halo mass, $M_{\rm vir}$, including peak quantities.  Models marked with a star are based on measures of halo maximum circular velocity, $V_{\rm max}$, including peak quantities, with the \citet{Lehmann:2017fy} model falling somewhere in between.   This paper focuses on the lack of models in the region marked with a ``?''. }
    \label{fig:venn_diagram}
\end{figure}

The goal of this paper is threefold: (i) to make explicit the tension between fitting galaxy clustering measurements and the evolution of the stellar mass function within the SHAM framework; (ii) to examine the successes and failures of SHAM models based on maximum circular velocity; and (iii) to evaluate mechanisms to alleviate the clustering crisis in models based on halo mass.  The structure of this paper is as follows: In \S \ref{sec:methods} we describe the SHAM models used in this work and our implementation of each.  In \S \ref{sec:clustering} we examine the galaxy clustering signal prediction of each model.  In \S \ref{sec:vpeak_miracle} we describe how it is that SHAM based on $V_{\rm peak}$ is able to fit clustering while models based on $M_{\rm peak}$ generally do not.  In \S \ref{sec:not_such_a_miracle} we discuss some pitfalls of $V_{\rm peak}$ SHAM and explicitly define the "crisis" referenced in the title of this paper.  \S \ref{sec:orphans}--\S \ref{sec:assem_bias} go through physically motivated alterations to $M_{\rm peak}$ based SHAM models which alleviate the clustering crisis to various degrees.  We provide a summary and discussion of the main conclusions of this paper in \S \ref{sec:discussion}.

The models and analysis in this paper utilize the framework and code base in {\tt Halotools} \citep{Hearin:2016tc}, an {\tt Astropy} \citep{TheAstropyCollaboration:2013cd} affiliated Python\footnote{\http{www.python.org}} package.  We also make available all of the code and data products necessary to reproduce the figures and analysis in this paper\footnote{\https{github.com/duncandc/Clustering\_Crisis}}.  Throughout, we scale all units by $h = H_0/[100~{\rm km/s}~{\rm Mpc}^{-1}]$ where appropriate, and we use $\log(x)$ to indicate the base 10 logarithm of $x$.

\section{Methods}
\label{sec:methods}

@@

SHAM, in its simplest form, assumes that the cumulative abundance of galaxies and haloes can be used to map galaxy properties uniquely onto (sub-)haloes.  Using stellar mass, $M_{*}$, and a halo mass proxy $\mathcal{M}$, SHAM assumes:
%
\begin{equation}
\label{eq:sham}
N(>M_*) = N(>\mathcal{M}) \implies M_{*} = f(\mathcal{M})
\end{equation} 
%
where $f(\mathcal{M})$ is some monotonically increasing function.  This maps the most massive galaxies to the most massive haloes.  The function $f(\mathcal{M})$ may be determined non-parametrically by matching the rank orders of galaxies and haloes.  We refer to this method of SHAM as ``rank order'' abundance matching.  However, we also consider another class of SHAM models where $f(\mathcal{M})$ is parametrized.  This introduces the possibility that the observed abundance of galaxies is not strictly matched; however, it is generally the goal of such models to preserve this characteristic. Here we use the umbrella term, SHAM, to refer to both the rank order abundance matching and parametrized SMHM relation methods.

@@

The simplest implementation of SHAM maps galaxies into (sub-)haloes by matching rank orders.  Given a set of galaxies, $M_*=\{M_{*i}\}$, and (sub-)haloes, $\mathcal{M}=\{\mathcal{M}_j\}$, of the same size and drawn from equivalent representative volumes, this method proceeds by calculating the ranks of each:
%
\begin{align}
n^{\rm gal} &= \mathcal{R}(M_*) \nonumber \\
n^{\rm halo} &= \mathcal{R}(\mathcal{M})
\end{align}
%
where the $\mathcal{R}(x)$ function returns the ordinal ranks.  Each galaxy is then assigned to a (sub-)halo with the equivalent rank, i.e. $n_i^{\rm gal}=n_j^{\rm halo}$.  We consider two rank order SHAM models in this paper, one based on $M_{\rm peak}$ and one on $V_{\rm peak}$.  As shorthand, we refer to these models as ``RM'' and ``RV'', respectively.

Given a set of galaxies and haloes, this mapping is deterministic.  However, the assumption of perfect rank ordering with respect to the cumulative distribution can be relaxed, allowing for stochasticity in the mapping and resulting in a probabilistic relation between stellar mass and halo mass, $P(M_*|\mathcal{M})$.  There are various methods used in the literature to add scatter to this relation while maintaining agreement with an observed stellar mass function.  One may attempt to deconvolve the stellar mass function from the scatter model such that, after solving for eq. (\ref{eq:sham}) with this deconvolved function and applying the scatter model, the new stellar mass function is consistent with the observed function \citep[e.g.][]{Behroozi:2010ja}.  Another option is to manually add scatter to the stellar masses used in abundance matching, re-ranking on the new values to perform the matching, and iteratively solving for a solution that results in the desired amount of scatter in the SMHM relation \citep[e.g.][]{Hearin:2013ok}.

We do not go through the additional step of adding scatter to our RM and RV models, as scatter in the SMHM relation generally \emph{decreases} the strength of the clustering signal for massive galaxies ($M_* > 10^{11} ~h^{-2} M_{\odot}$) with little effect at lower masses \citep{Tinker:2016vu}, and our work focuses on the \emph{decrement} in galaxy clustering signal for samples of less massive galaxies in SHAM models.

@@

A distinct method from the rank order SHAM method is to parametrize the SMHM relation, fitting for the parameters which result in a stellar mass function that is consistent with an observed function after populating a dark matter simulation.  The first model of this type we consider is by \cite{Moster:2013ab}. In this model, the SMHM relation is parametrized as a function of halo mass and redshift to account for evolution.  Furthermore, it is assumed that the stellar mass of satellite galaxies is determined by the mass of sub-haloes at the time of accretion into a more massive host-halo for the first time.  The functional form is given by:
%
\begin{equation}
\label{eq:moster_smhm}
\frac{\langle M_{*} | \mathcal{M}\rangle_{\rm med}(a)}{\mathcal{M}} = 2N\left[ \left(\frac{\mathcal{M}}{{M}_1} \right)^{-\beta} + \left( \frac{\mathcal{M}}{{M}_1} \right)^{\gamma} \right]^{-1}
\end{equation}
%
where $\langle\rangle_{\rm med}$ indicates the median.  The evolution of the parameters is given by:
%
\begin{align}
\label{eq:smhm_params}
\log[{M}_1(a)]& = {M}_{10} + {M}_{11}(1-a) \\
N(a) &= N_{10} + N_{11}(1-a) \\
\beta(a) &= \beta_{10} + \beta_{11}(1-a) \\ 
\gamma(a) &=  \gamma_{10} + \gamma_{11}(1-a)
\end{align}
%
where the scale factor is either the instantaneous scale factor, or the one at the time of accretion:
%
\begin{equation}
a = 
\begin{cases}
    a      & \quad \text{if host-halo}\\
    a_{\rm acc}  & \quad \text{if sub-halo} \\
  \end{cases}
\end{equation}
%
and $\mathcal{M}$ is the instantaneous mass or the mass at accretion, $M_{\rm acc}$, for sub-haloes.  Scatter in the SMHM relation at fixed halo mass is modelled as a fixed log-normal with $\sigma_{\log(M*)} \approx 0.18$.

\citet{Moster:2013ab} find the best fit parameters for equation (\ref{eq:moster_smhm}) that reproduce the stellar mass function at various redshifts between $z=0$ and $z \sim 4$.  They also show that the implied star-formation rates of galaxies given the growth history of the (sub-)haloes in their simulation is consistent with the cosmic star-formation rate density.  The parameters were constrained using the Millennium simulation with (sub-)haloes defined as spherical over-densities (SO) with mean internal density 200 times the critical density of the universe, $M_{200c}$.   As shorthand, we will refer to this model as ``M13''.  The parameter values used in this paper are taken directly from \citet{Moster:2013ab} and listed in table 2.

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{figures/stellar_mass_function.pdf}
    \caption{The stellar mass function of a mock realization of each model is plotted (lines).  For comparison we show the stellar mass function in SDSS as measured by \citet{Li:2009kh} as points with error bars.  The lower panel shows the residuals with respect to the triple Schechter fit from \citet{Li:2009kh}. As an example, we show the $\pm 1 \sigma$ error on the prediction for the RV model as the shaded green region, calculated by re-populating the simulation with the model $50$ times. While not shown, the uncertainty on the other model predictions are similar.}
    \label{fig:phi}
\end{figure}

(Sub-)haloes are found using the phase-space halo finder {\tt ROCKSTAR} \citep{Behroozi:2013cn}, which uses adaptive, hierarchical refinement of friends-of-friends groups in six phase-space dimensions and one time dimension, and tracked over time using the {\tt Consistent Trees} algorithm \citep{Behroozi:2013dz}.  As demonstrated in \cite{Knebe:2011jc, Knebe:2013bp}, this results in a very robust tracking of (sub-)haloes \citep[also see][]{vandenBosch:2016ky}. Haloes in this catalogue are defined to be spherical volumes centred on a local density peak (SO hereafter), such that the average density inside the sphere is $\bar{\rho}_{\rm h}(z) = \Delta_{\rm vir}(z)\rho_{\rm m}(z)$.  Here $\rho_{\rm m}(z) = \Omega_{\rm m}(z) \rho_{\rm crit}(z)$, where $\rho_{\rm crit}(z) = 3H(z)^2/8\pi G$ is the critical energy density of the Universe, and $\Delta_{\rm vir}(z)$ is given by a fitting function \citep{Bryan:1998cc}:
%
\begin{equation}
\Delta_{\rm vir}(z) = \left[ 18\pi^2 - 82\Omega_{\Lambda}(z) - 39\Omega_{\Lambda}^2(z) \right]\Omega_m^{-1}
\end{equation}
%
For the Bolshoi cosmology, $\Delta_{\rm vir}(z=0)\simeq 360$.  The radius of each such sphere defines the virial radius $R_{\rm vir}$ of the halo, which is related to the mass of the halo via $M_{\rm vir} = (4/3)\pi R_{\rm vir}^3\bar{\rho}_{\rm h}$. Additionally, sub-haloes in this catalogue are distinct, self-bound structures whose centres are found within the virial radius of a more massive host-halo.  For each (sub-)halo, the maximum circular velocity is defined as: $V_{\rm max} \equiv {\rm Max}[GM(<r)/r]$, where $M(<r)$ is the mass enclosed within a distance r of the (sub-)halo centre.
